### 2020-07-27 TCS_v1/06-examples.tex
pulizia e riorganizzazione

Hereafter, we shall often omit topmost restrictions but shall take into account only transitions whose labels are complete, i.e. they do not contain 
virtual links and start/end with the $\silent$ action symbol.

\subsection{Labelled transition system}
This example is inspired by the example in~\cite{BEMR11}, where
a deterministic transition system is coded in the Reaction System
framework.
Here we consider the minimal deterministic transition system in Figure~\ref{fig:lts}.
%following the reaction system approach, to code it in the reaction system framework, we have to add
%extra labels, as in Figure~\ref{fig:ndlts2}.
\begin{figure}
\[
\xymatrix{
*+[o][F-]{q} \ar@/ ^/[r]^a \ar@(l,u)[]^b &
*+[o][F-]{w} \ar@/^/[l]^b \ar@(r,u)[]_a
}
\]
\caption{Minimal deterministic labelled transition system.}
\label{fig:lts}
\end{figure}
\noindent
In our $c\CNA$ encoding, $q$, $w$, $a$, and $b$ became the $entities$, the context can only provide $a$ and $b$ and the reaction rules are as follows:\\
\[
\begin{array}{cl@{\hspace{1cm}} cl}
1 \ &(\{q,a\},\{w,b\},\{q\}) & 2\ &(\{q,b\},\{w,a\},\{w\})\\
3 \  &(\{w,a\},\{q,b\},\{q\}) & 4\ &(\{w,b\},\{q,a\},\{w\})\\ 
\end{array}
\]
\paragraph{Encoding of the rules}
The encoding of the rules for reactions is  given in a parametric way:
\[
\begin{array}{lcl}
P_n(q,b,w,a,q) & \defeq & \pi_n(q,b,w,a,q).P_n(q,b,w,a,q)  \\
			&& +\\
			&& \sum_{x\in\{\overline{q},\overline{b},w,a\}}\ \pi'_n(x).P_n(q,b,w,a,q)\\
%			&& +\ \pi'(\overline{b}).P_n(q,b,w,a,q)\\
%			&& +\ \pi'(w).P_n(q,b,w,a,q)\\ 
%			&& +\ \pi'(a).P_n(q,b,w,a,q)\\
			\end{array}
			\]
			\noindent
			where
			\[
\begin{array}{lcl}
\pi_n(q,b,w,a,q) &\defeq & \startchain{r_n}\chainedlink{q_i}{\noact}\chainedlink{\noact}{q_o}
                                                    \chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}
                                                     \chainedlink{\overline{w}_i}{\noact}\chainedlink{\noact}{\overline{w}_o}
					        \chainedlink{\overline{a}_i}{\noact}\chainedlink{\noact}{\overline{a}_o}
					        \chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n}
					        \chainedlink{\widetilde{q}_i}{\noact}\chainedlink{\noact}{\widetilde{q}_o}
			\chainend{p_{n+1}}\\[6pt]
\pi'_n(x) &\defeq &  \startchain{r_n}\chainedlink{x_i}{\noact}\chainedlink{\noact}{x_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}} 
% \ + \   \startchain{r_n}\chainedlink{w_i}{\noact}\chainedlink{\noact}{w_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n(q,b,w,a,q) \\
%			&&+\\
%			&& \startchain{r_n}\chainedlink{\overline{b}_i}{\noact}\chainedlink{\noact}{\overline{b}_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n}\chainend{p_{n+1}}.P_n(q,b,w,a,q) \ + \  \startchain{r_n}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n(q,b,w,a,q) 
\end{array}
\]
\noindent
Then we have 
\[
\begin{array}{lcl @{\hspace{0.5cm}}lcl  }
P_1 & \defeq & P_1(q,a,w,b,q) & P_3 & \defeq & P_3(w,a,q,b,w)  \\
P_2 & \defeq & P_2(q,b,w,a,w) &P_4 & \defeq & P_4(w,b,q,a,q) 
\end{array}
\]
\noindent
and we put, as usual,
%$r_1 = \silent$, 
$r_5 = \mathit{cxt}$ and  $p_5 = \silent$.

@@

Now, we assume that we have an initial configuration containing the entities $q$ and $b$:
% (with $\widetilde{q},\widetilde{w},\widetilde{a},\widetilde{b}$ grouping all the possibile decorations for these names):
\[
\begin{array}{lcl}
\mathit{Sys}  \defeq \ %\restrict{r1,r2,r3,r4,\widetilde{q},\widetilde{w},\widetilde{a},\widetilde{b}}
 I\mid P_q \mid \overline{P}_w  \mid  \overline{P}_a  \mid P_b  \mid  P_1  \mid  P_2  \mid  P_3  \mid P_4  \mid \mathit{Cxt}.
\end{array}
\] 
Then, only the second reaction can be applied, and the transition carries the complete label $\upsilon$ below
{\tiny
\[
%\begin{array}{l}
%\upsilon = \\
%\restrict{r1,r2,r3,r4,\widetilde{q},\widetilde{w},\widetilde{a},\widetilde{b}}\\
 \startchain{\silent}
 \chainedlink{r_1}{r_1}
 \chainedlink{\overline{a}_i}{\bf  \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
 \chainedlink{r_2}{r_2}
 \chainedlink{q_i}{\bf q_i}\chainedlink{\bf q_o}{q_o}
 \chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_o}{b_o}
 \chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}
 \chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
 \chainedlink{r_3}{r_3}
 \chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
 \chainedlink{r_4}{r_4}
 \chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}
 \chainedlink{\mathit{cxt}}{\mathit{cxt}}
 \chainedlink{\widehat{a}_i}{\bf \widehat{a}_i}\chainedlink{\bf \widehat{a}_o}{\widehat{a}_o}
 \chainedlink{\underline{b}_i}{\bf \underline{b}_i}\chainedlink{\bf \underline{b}_o}{\underline{b}_o}
 \chainedlink{p_1}{p_1}
 \chainedlink{p_2}{p_2}
 \chainedlink{\widetilde{w}_i}{\bf \widetilde{w}_i}\chainedlink{\bf \widetilde{w}_o}{\widetilde{w}_o}
 \chainedlink{p_3}{p_3}
 \chainedlink{p_4}{p_4}
 \chainend{\silent}
%\end{array}
\]}
The parts in bold are provided by the entity processes, the other parts are provided by the processes encoding the reactions and by the process encoding the context (starting at $\mathit{cxt}$ and ended at $p_1$.
In the label we can read that the rules $1$ and $4$ have been not executed because the entity $a$ is absent, 
the rule $3$ has been not applied because the entity $w$ is absent, then only rule $2$ has been applied, and it has produced the entity $w$. Also,  the context provides entity $a$, that will be available in the next state, and not the entity $b$.
Now, to let the label more readable, we show the result of the application of the function $flat(\cdot)$ to it:
\[
r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~\mathit{cxt}~\widehat{a}~\underline{b}~p_1~p_2~\widetilde{w}~p_3~p_4\]

@@

Here, the entities $T$ and $T'$ that can be produced by the reactions and that are not provided by the context:
\[
\begin{array}{lcllcl}
P (T,\widetilde{T})&\defeq & \sum_{h=0}^1 (\startchain{T_i}\chainedlink{T_o}{\noact}\chainend{\noact})^h\ \link{\widetilde{T}_i}{\widetilde{T}_o}.P (T,\widetilde{T})
& +  && \link{T_i}{T_o}.P (\overline{T},\widetilde{T}) \\
%\overline{P} (q)&\defeq &  \sum_{h=0}^1 (\startchain{\overline{T}_i}\chainedlink{\overline{T}_o}{\noact}\chainend{\noact})^h\link{\widetilde{T}_i}{\widetilde{T}_o}.P(q)\\
%&& + & && +\\
%&& \link{T_i}{T_o}.\overline{P}(T) &
%&& \link{\overline{T}_i}{\overline{T}_o}.\overline{P}(T)\\
\end{array}
\]
Then, we have
\[
\begin{array}{lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl}
P_T & \defeq & P(T,\widetilde{T})& \overline{P}_T & \defeq & P(\overline{T},\widetilde{T}) & P_{T'}& \defeq & P(T',\widetilde{T'}) & \\
\overline{P}_{T',\widetilde{T'}}& \defeq & P(\overline{T'},\widetilde{T'})\\\
\end{array}
\]
The entities that can be produced by the reactions and that can be  provided by the context are as follows:
\[
\begin{array}{lcl @{\hspace{1cm}}lcl}
P(C,\widehat{C},\underline{C},\widetilde{C}) & \defeq &  \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\widehat{C}_i}{\widehat{C}_o} (\startchain{\noact}\chainedlink{\noact}{\widetilde{C}_i}\chainend{\widetilde{C}_o})^h.   P(C,\widehat{C},\underline{C},\widetilde{C})\\
&&+\\
&& \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\underline{C}_i}{\underline{C}_o}.P(\overline{C},\widehat{C},\underline{C},\widetilde{C})\\
&&+\\
 && \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\underline{C}_i}{\underline{C}_o} \startchain{\noact}\chainedlink{\noact}{\widehat{C}_i}\chainend{\widehat{C}_o}.   P(C,\widehat{C},\underline{C},\widetilde{C}) \\
 %&  \overline{E} (a)&\defeq & \sum_{h=1}^2 (\startchain{\overline{a}_i}\chainedlink{\overline{a}_o}{\noact}\chainend{\noact})^h\ \link{\widehat{a}_i}{\widehat{a}_o}.E(a)\\
%&&  &  && +\\
%&& 
%&
%&& \sum_{h=1}^2 (\startchain{\overline{a}_i}\chainedlink{\overline{a}_o}{\noact}\chainend{\noact})^h\ \link{\underline{a}_i}{\underline{a}_o}.\overline{E}(q)\\
\end{array}
\]
Then, we have
\[
\begin{array}{lcl @{\hspace{1cm}}lcl@{\hspace{0.4cm}}lcl@{\hspace{0.4cm}}lcl}
P_C & \defeq &  P(C,\widehat{C},\underline{C},\widetilde{C})&  P_G& \defeq &   P(G,\widehat{G},\underline{G},\widetilde{G}) \\[6pt]
\overline{P}_C & \defeq &  P(\overline{C},\widehat{C},\underline{C},\widetilde{C}) & 
\overline{P}_G& \defeq &   P(\overline{G},\widehat{G},\underline{G},\widetilde{G})\\
\end{array}
\]

@@

Now, to show a possible composition of a transition label, we assume a system where only the entities $a$, $T'$, and $G$ are present:
\[
\begin{array}{lcl}
\mathit{Sys} &\defeq &%\restrict{\widetilde{C},\widetilde{G},\widetilde{F}, \widetilde{T},\widetilde{T'}}(
I\mid P_a\mid \overline{P}_C\mid P_G\mid \overline{P}_F \mid \overline{P}_T \mid P_{T'} \mid P_1 \mid P_2 \mid P_3 \mid P_4 \mid \mathit{Cxt}
\end{array}
\]
In the above configuration, reactions $1$ and $4$ can be applied, and also we assume that the context will provide the entity $F$, that will be available in the target configuration.
Instead of showing the complete transition label, we give its flattened version obtained by applying the function $\mathit{flat}(\cdot)$:
%The transition label appears as:
\[
%\restrict{r1,r2,r3,r4,a,\widetilde{C},\widetilde{G},\widetilde{F},\widetilde{T},\widetilde{T'}}\\
% \startchain{\silent}\chainedlink{a_i}{\bf  a_i}\chainedlink{\bf a_o}{a_o}
% \chainedlink{G_i}{\bf  G_i}\chainedlink{\bf G_o}{G_o}
% \chainedlink{\overline{C}_i}{\bf \overline{C}_i}\chainedlink{\bf \overline{C}_o}{\overline{C}_o}
% \chainedlink{r_2}{r_2}\chainedlink{G_i}{\bf G_i}\chainedlink{\bf G_o}{G_o}
%\chainedlink{r_3}{r_3}\chainedlink{T'_i}{\bf T'_i}\chainedlink{\bf T'_o}{T'_o}\chainedlink{r_4}{r_4}\chainedlink{T'_i}{\bf T'_i}\chainedlink{\bf T'_o}{T'_o}\chainedlink{\overline{F}_i}{\bf \overline{F}_i}\chainedlink{\bf \overline{F}_o}{\overline{F}_o}\chainedlink{cxt}{cxt}
%\chainedlink{\underline{C}_i}{\bf \underline{C}_i}\chainedlink{\bf \underline{C}_o}{\underline{C}_o}
%\chainedlink{\underline{G}_i}{\bf \underline{G}_i}\chainedlink{\bf \underline{G}_o}{\underline{G}_o}
%\chainedlink{\widehat{F}_i}{\bf \widehat{F}_i}\chainedlink{\bf \widehat{F}_o}{\widehat{F}_o}\chainedlink{p_1}{p_1} 
% \chainedlink{\widetilde{T}_i}{\bf \widetilde{T}_i}\chainedlink{\bf \widetilde{T}_o}{\widetilde{T}_o}
%  \chainedlink{p_2}{p_2}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4} \chainedlink{\widetilde{G}_i}{\bf \widetilde{G}_i}\chainedlink{\bf \widetilde{G}_o}{\widetilde{G}_o} \chainend{\silent}
r_1~ a~G~\overline{C}~ r_2~ G~ r_3~T'~r_4~T'~\overline{F}~cxt~\underline{C}~\underline{G}~\widehat{F}~p_1~\widetilde{T}~p_2~p_3~p_4~\widetilde{G}
\]
The original label can then be reconstructed just applying the function $\mathit{unflat}(\cdot)$ to the string above.

>>>>

Hereafter, we shall often omit topmost restrictions but shall take into account only transitions whose labels are complete chains, i.e. they do not contain 
virtual links and start/end with the $\silent$ action symbol.

\subsection{Labelled transition system}
This example is inspired by the example in~\cite{BEMR11}, where
a deterministic transition system is coded in the Reaction System
framework.
Here we consider the minimal deterministic transition system in Figure~\ref{fig:lts}.
%following the reaction system approach, to code it in the reaction system framework, we have to add
%extra labels, as in Figure~\ref{fig:ndlts2}.
\begin{figure}
\[
\xymatrix{
*+[o][F-]{q} \ar@/ ^/[r]^a \ar@(l,u)[]^b &
*+[o][F-]{w} \ar@/^/[l]^b \ar@(r,u)[]_a
}
\]
\caption{Minimal deterministic labelled transition system.}
\label{fig:lts}
\end{figure}
\noindent
In our $c\CNA$ encoding, $q$, $w$, $a$, and $b$ became the $entities$, the context can only provide $a$ and $b$ and the reaction rules are as follows:\\
\[
\begin{array}{cl@{\hspace{1cm}} cl}
1 \ &(\{q,a\},\{w,b\},\{q\}) & 2\ &(\{q,b\},\{w,a\},\{w\})\\
3 \  &(\{w,a\},\{q,b\},\{q\}) & 4\ &(\{w,b\},\{q,a\},\{w\})\\ 
\end{array}
\]
\paragraph{Encoding of the rules}
The encoding of the rules for reactions is  given in a parametric way:
\[
\begin{array}{lcl}
P_n(q,b,w,a,q) & \defeq & \pi_n(q,b,w,a,q).P_n(q,b,w,a,q)  \\
			&& +\\
			&& \sum_{x\in\{\overline{q},\overline{b},w,a\}}\ \pi'_n(x).P_n(q,b,w,a,q)\\
%			&& +\ \pi'(\overline{b}).P_n(q,b,w,a,q)\\
%			&& +\ \pi'(w).P_n(q,b,w,a,q)\\ 
%			&& +\ \pi'(a).P_n(q,b,w,a,q)\\
			\end{array}
			\]
			\noindent
			where
			\[
\begin{array}{lcl}
\pi_n(q,b,w,a,q) &\defeq & \startchain{r_n}\chainedlink{q_i}{\noact}\chainedlink{\noact}{q_o}
                                                    \chainedlink{b_i}{\noact}\chainedlink{\noact}{b_o}
                                                     \chainedlink{\overline{w}_i}{\noact}\chainedlink{\noact}{\overline{w}_o}
					        \chainedlink{\overline{a}_i}{\noact}\chainedlink{\noact}{\overline{a}_o}
					        \chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n}
					        \chainedlink{\widetilde{q}_i}{\noact}\chainedlink{\noact}{\widetilde{q}_o}
			\chainend{p_{n+1}}\\[6pt]
\pi'_n(x) &\defeq &  \startchain{r_n}\chainedlink{x_i}{\noact}\chainedlink{\noact}{x_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}} 
% \ + \   \startchain{r_n}\chainedlink{w_i}{\noact}\chainedlink{\noact}{w_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n(q,b,w,a,q) \\
%			&&+\\
%			&& \startchain{r_n}\chainedlink{\overline{b}_i}{\noact}\chainedlink{\noact}{\overline{b}_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n}\chainend{p_{n+1}}.P_n(q,b,w,a,q) \ + \  \startchain{r_n}\chainedlink{a_i}{\noact}\chainedlink{\noact}{a_o}\chainedlink{r_{n+1}}{\noact}\chainedlink{\noact}{p_n} \chainend{p_{n+1}}.P_n(q,b,w,a,q) 
\end{array}
\]

Then we have 
\[
\begin{array}{lcl @{\hspace{1cm}}lcl  }
P_1 & \defeq & P_1(q,a,w,b,q) & P_3 & \defeq & P_3(w,a,q,b,w)  \\
P_2 & \defeq & P_2(q,b,w,a,w) &P_4 & \defeq & P_4(w,b,q,a,q) 
\end{array}
\]
and we put, as usual,
%$r_1 = \silent$, 
$r_5 = \mathit{cxt}$ and  $p_5 = \silent$.

@@

Now, we assume that we have an initial configuration containing the entities $q$ and $b$:
% (with $\widetilde{q},\widetilde{w},\widetilde{a},\widetilde{b}$ grouping all the possibile decorations for these names):
\[
\begin{array}{lcl}
\mathit{Sys}  \defeq \ %\restrict{r1,r2,r3,r4,\widetilde{q},\widetilde{w},\widetilde{a},\widetilde{b}}
 I\mid P_q \mid \overline{P}_w  \mid  \overline{P}_a  \mid P_b  \mid  P_1  \mid  P_2  \mid  P_3  \mid P_4  \mid \mathit{Cxt}.
\end{array}
\]

Then, only the second reaction can be applied, and the transition carries the complete label $\upsilon$ below
{\tiny
\[
%\begin{array}{l}
%\upsilon = \\
%\restrict{r1,r2,r3,r4,\widetilde{q},\widetilde{w},\widetilde{a},\widetilde{b}}\\
 \startchain{\silent}
 \chainedlink{r_1}{r_1}
 \chainedlink{\overline{a}_i}{\bf  \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
 \chainedlink{r_2}{r_2}
 \chainedlink{q_i}{\bf q_i}\chainedlink{\bf q_o}{q_o}
 \chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_o}{b_o}
 \chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}
 \chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
 \chainedlink{r_3}{r_3}
 \chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
 \chainedlink{r_4}{r_4}
 \chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}
 \chainedlink{\mathit{cxt}}{\mathit{cxt}}
 \chainedlink{\widehat{a}_i}{\bf \widehat{a}_i}\chainedlink{\bf \widehat{a}_o}{\widehat{a}_o}
 \chainedlink{\underline{b}_i}{\bf \underline{b}_i}\chainedlink{\bf \underline{b}_o}{\underline{b}_o}
 \chainedlink{p_1}{p_1}
 \chainedlink{p_2}{p_2}
 \chainedlink{\widetilde{w}_i}{\bf \widetilde{w}_i}\chainedlink{\bf \widetilde{w}_o}{\widetilde{w}_o}
 \chainedlink{p_3}{p_3}
 \chainedlink{p_4}{p_4}
 \chainend{\silent}
%\end{array}
\]}
The parts in bold are provided by the entity processes, the other parts are provided by the processes encoding the reactions and by the process encoding the context (starting at $\mathit{cxt}$ and ended at $p_1$.
In the label we can read that the rules $1$ and $4$ have been not executed because the entity $a$ is absent, 
the rule $3$ has been not applied because the entity $w$ is absent, then only rule $2$ has been applied, and it has produced the entity $w$. Also,  the context provides entity $a$, that will be available in the next state, and not the entity $b$.
Now, to let the label more readable, we show the result of the application of the function $flat(\cdot)$ to it:
\[
r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~\mathit{cxt}~\widehat{a}~\underline{b}~p_1~p_2~\widetilde{w}~p_3~p_4 .
\]

@@

Here, the entities $T$ and $T'$ that can be produced by the reactions and that are not provided by the context:
\[
\begin{array}{lcllcl}
P (T,\widetilde{T})&\defeq & \sum_{h=0}^1 (\startchain{T_i}\chainedlink{T_o}{\noact}\chainend{\noact})^h\ \link{\widetilde{T}_i}{\widetilde{T}_o}.P (T,\widetilde{T})
& +  && \link{T_i}{T_o}.P (\overline{T},\widetilde{T}) \\
%\overline{P} (q)&\defeq &  \sum_{h=0}^1 (\startchain{\overline{T}_i}\chainedlink{\overline{T}_o}{\noact}\chainend{\noact})^h\link{\widetilde{T}_i}{\widetilde{T}_o}.P(q)\\
%&& + & && +\\
%&& \link{T_i}{T_o}.\overline{P}(T) &
%&& \link{\overline{T}_i}{\overline{T}_o}.\overline{P}(T)\\
\end{array}
\]

Then, we have
\[
\begin{array}{lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl @{\hspace{0.5cm}}lcl}
P_T & \defeq & P(T,\widetilde{T})& \overline{P}_T & \defeq & P(\overline{T},\widetilde{T}) & P_{T'}& \defeq & P(T',\widetilde{T'}) & \\
\overline{P}_{T',\widetilde{T'}}& \defeq & P(\overline{T'},\widetilde{T'})
\end{array}
\]

The entities that can be produced by the reactions and that can be  provided by the context are as follows:
\[
\begin{array}{lcl @{\hspace{1cm}}lcl}
P(C,\widehat{C},\underline{C},\widetilde{C}) & \defeq &  \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\widehat{C}_i}{\widehat{C}_o} (\startchain{\noact}\chainedlink{\noact}{\widetilde{C}_i}\chainend{\widetilde{C}_o})^h.   P(C,\widehat{C},\underline{C},\widetilde{C})\\
&&+\\
&& \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\underline{C}_i}{\underline{C}_o}.P(\overline{C},\widehat{C},\underline{C},\widetilde{C})\\
&&+\\
 && \sum_{h=0}^1 (\startchain{C_i}\chainedlink{C_o}{\noact}\chainend{\noact})^h\ \link{\underline{C}_i}{\underline{C}_o} \startchain{\noact}\chainedlink{\noact}{\widehat{C}_i}\chainend{\widehat{C}_o}.   P(C,\widehat{C},\underline{C},\widetilde{C}) \\
 %&  \overline{E} (a)&\defeq & \sum_{h=1}^2 (\startchain{\overline{a}_i}\chainedlink{\overline{a}_o}{\noact}\chainend{\noact})^h\ \link{\widehat{a}_i}{\widehat{a}_o}.E(a)\\
%&&  &  && +\\
%&& 
%&
%&& \sum_{h=1}^2 (\startchain{\overline{a}_i}\chainedlink{\overline{a}_o}{\noact}\chainend{\noact})^h\ \link{\underline{a}_i}{\underline{a}_o}.\overline{E}(q)\\
\end{array}
\]

Then, we have
\[
\begin{array}{lcl @{\hspace{1cm}}lcl@{\hspace{0.4cm}}lcl@{\hspace{0.4cm}}lcl}
P_C & \defeq &  P(C,\widehat{C},\underline{C},\widetilde{C})&  P_G& \defeq &   P(G,\widehat{G},\underline{G},\widetilde{G}) \\[6pt]
\overline{P}_C & \defeq &  P(\overline{C},\widehat{C},\underline{C},\widetilde{C}) & 
\overline{P}_G& \defeq &   P(\overline{G},\widehat{G},\underline{G},\widetilde{G})\\
\end{array}
\]

@@

Now, to show a possible composition of a transition label, we assume a system where only the entities $a$, $T'$, and $G$ are present:
\[
\begin{array}{lcl}
\mathit{Sys} &\defeq &%\restrict{\widetilde{C},\widetilde{G},\widetilde{F}, \widetilde{T},\widetilde{T'}}(
I\mid P_a\mid \overline{P}_C\mid P_G\mid \overline{P}_F \mid \overline{P}_T \mid P_{T'} \mid P_1 \mid P_2 \mid P_3 \mid P_4 \mid \mathit{Cxt}
\end{array}
\]

In the above configuration, reactions $1$ and $4$ can be applied, and also we assume that the context will provide the entity $F$, that will be available in the target configuration.
Instead of showing the complete transition label, we give its flattened version obtained by applying the function $\mathit{flat}(\cdot)$:
%The transition label appears as:
\[
%\restrict{r1,r2,r3,r4,a,\widetilde{C},\widetilde{G},\widetilde{F},\widetilde{T},\widetilde{T'}}\\
% \startchain{\silent}\chainedlink{a_i}{\bf  a_i}\chainedlink{\bf a_o}{a_o}
% \chainedlink{G_i}{\bf  G_i}\chainedlink{\bf G_o}{G_o}
% \chainedlink{\overline{C}_i}{\bf \overline{C}_i}\chainedlink{\bf \overline{C}_o}{\overline{C}_o}
% \chainedlink{r_2}{r_2}\chainedlink{G_i}{\bf G_i}\chainedlink{\bf G_o}{G_o}
%\chainedlink{r_3}{r_3}\chainedlink{T'_i}{\bf T'_i}\chainedlink{\bf T'_o}{T'_o}\chainedlink{r_4}{r_4}\chainedlink{T'_i}{\bf T'_i}\chainedlink{\bf T'_o}{T'_o}\chainedlink{\overline{F}_i}{\bf \overline{F}_i}\chainedlink{\bf \overline{F}_o}{\overline{F}_o}\chainedlink{cxt}{cxt}
%\chainedlink{\underline{C}_i}{\bf \underline{C}_i}\chainedlink{\bf \underline{C}_o}{\underline{C}_o}
%\chainedlink{\underline{G}_i}{\bf \underline{G}_i}\chainedlink{\bf \underline{G}_o}{\underline{G}_o}
%\chainedlink{\widehat{F}_i}{\bf \widehat{F}_i}\chainedlink{\bf \widehat{F}_o}{\widehat{F}_o}\chainedlink{p_1}{p_1} 
% \chainedlink{\widetilde{T}_i}{\bf \widetilde{T}_i}\chainedlink{\bf \widetilde{T}_o}{\widetilde{T}_o}
%  \chainedlink{p_2}{p_2}\chainedlink{p_3}{p_3}\chainedlink{p_4}{p_4} \chainedlink{\widetilde{G}_i}{\bf \widetilde{G}_i}\chainedlink{\bf \widetilde{G}_o}{\widetilde{G}_o} \chainend{\silent}
r_1~ a~G~\overline{C}~ r_2~ G~ r_3~T'~r_4~T'~\overline{F}~cxt~\underline{C}~\underline{G}~\widehat{F}~p_1~\widetilde{T}~p_2~p_3~p_4~\widetilde{G} .
\]

The original label can then be reconstructed just applying the function $\mathit{unflat}(\cdot)$ to the string above.

### 2020-07-27 TCS_v1/07-discussion.tex
pulizia e riorganizzazione

\subsection{Mutating entities} 
In RS, when an entity is present, it can potentially be involved in each reactions where it is required.
With a few more lines of code, in c\CNA~it is possible to describe the behaviour of a mutation of 
an entity, in a way that the mutated version of the entity can take 
part to only a subset of the rules requiring  the  \emph{normal version} of the entity.
For example, let us assume that entity $s1$ is consumed by reactions $a1$ and $a2$.
Reaction $a1$ produces also $s1$ if $s2$ is present, otherwise $a1$ produces a mutated version of $s1$, 
say $s1'$.
When $s1'$ is produced, reaction $a2$ behaves in the same way as if $s1$ would be absent, whereas $a2$
recognises the presence of $s1'$ and behaves in the same way as if $s1$ would be present.
Technically, in both cases it is enough to add one more non 
deterministic choice in the code of $P_{a1}$ and $P_{a2}$.

\subsection{Communicating Reaction Systems}
We sketch how it is possible to program two RSs encodings, in a way that
the entities that usually come from context of one RS will be provided instead from the other RS.

>>>>

\subsection{Mutating entities} 
In RS, when an entity is present, it can potentially be involved in each reactions where it is required.
With a few more lines of code, in c\CNA~it is possible to describe the behaviour of a mutation of 
an entity, in a way that the mutated version of the entity can take 
part to only a subset of the rules requiring  the  \emph{normal version} of the entity.
For example, let us assume that entity $s1$ is consumed by reactions $a1$ and $a2$.
Reaction $a1$ produces also $s1$ if $s2$ is present, otherwise $a1$ produces a mutated version of $s1$, 
say $s1'$.
When $s1'$ is produced, reaction $a2$ behaves in the same way as if $s1$ would be absent, whereas $a2$
recognises the presence of $s1'$ and behaves in the same way as if $s1$ would be present.
Technically, in both cases it is enough to add one more nondeterministic choice in the code of $P_{a1}$ and $P_{a2}$.

\subsection{Communicating Reaction Systems}
We sketch how it is possible to program two RSs encodings, in a way that
the entities that usually come from the context of one RS will be provided instead from the other RS.
